---
title: "Covid-19 in African Countries"
author: "Jackeline Romio"
date: "13/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#I will analyze the trends about covid-19 deaths cases and imunization from april 2020 to march 2022, in African countries.
```{r, echo=FALSE}
#install.packages("devtools")
#devtools::install_github("RamiKrispin/coronavirus")
library(ggplot2)
library(patchwork)
library(lubridate)
library(zoo)
library(dplyr)
library(coronavirus)
#update_dataset() #extint
```

```{r}
data(coronavirus)
head(coronavirus)
```
#Total de casos período 
#Cases confirmed
#Eastern Africa 1 

```{r}
weekly=coronavirus%>%filter(type=="confirmed", country %in% c("Ethiopia", "Tanzania", "Kenya", "Uganda", "Mozambique", "Madagascar", "Malawi", "Zambia", "Somalia", "Zimbabwe"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly confirmed cases")+theme_minimal()
```

#Eastern Africa 2

```{r}
weekly=coronavirus%>%filter(type=="confirmed", country %in% c("Rwanda", "Burundi", "South Sudan", "Eritrea", "Mauritius", "Djibouti", "Comoros", "Seychelles"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly confirmed cases")+theme_minimal()
```


#Middle Africa

```{r}
weekly=coronavirus%>%filter(type=="confirmed", country %in% c("DR Congo", "Angola", "Cameroon", "Chad", "Congo", "Central African Republic", "Gabon", "Equatorial Guinea", "Sao Tome and Principe"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly confirmed cases")+theme_minimal()
```


#Northern Africa

```{r}
weekly=coronavirus%>%filter(type=="confirmed", country %in% c("Egypt", "Sudan", "Algeria", "Morocco", "Tunisia", "Libya"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly confirmed cases")+theme_minimal()
```

#Southern Africa

```{r}
weekly=coronavirus%>%filter(type=="confirmed", country %in% c("South Africa", "Botswana", "Namibia", "Lesotho", "Eswatini"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly confirmed cases")+theme_minimal()
```


#Western Africa 1

```{r}
weekly=coronavirus%>%filter(type=="confirmed", country %in% c("Nigeria", "Ghana", "Côte d'Ivoire", "Niger", "Mali", "Burkina Faso", "Senegal", "Guinea"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly confirmed cases")+theme_minimal()
```

#Western Africa 2

```{r}
weekly=coronavirus%>%filter(type=="confirmed", country %in% c("Benin", "Togo", "Sierra Leone", "Liberia", "Mauritania", "Gambia", "Guinea-Bissau", "Cabo Verde"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly confirmed cases")+theme_minimal()
```







#Deaths


#Eastern African 1


```{r}
weekly=coronavirus%>%filter(type=="death", country %in% c("Ethiopia", "Tanzania", "Kenya", "Uganda", "Mozambique", "Madagascar", "Zambia", "Somalia", "Zimbabwe"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly deaths")+theme_minimal()
```



Eastern Africa 2

```{r}
weekly=coronavirus%>%filter(type=="death", country %in% c("Rwanda", "Burundi", "South Sudan", "Eritrea", "Mauritius", "Djibouti", "Comoros", "Seychelles"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly deaths")+theme_minimal()
```


Middle Africa

```{r}
weekly=coronavirus%>%filter(type=="death", country %in% c("DR Congo", "Angola", "Cameroon", "Chad", "Congo", "Central African Republic", "Gabon", "Equatorial Guinea", "Sao Tome and Principe"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly deaths")+theme_minimal()
```


#Northern Africa

```{r}
weekly=coronavirus%>%filter(type=="death", country %in% c("Egypt", "Sudan", "Algeria", "Morocco", "Tunisia", "Libya"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly deaths")+theme_minimal()
```

#Southern Africa

```{r}
weekly=coronavirus%>%filter(type=="death", country %in% c("South Africa", "Botswana", "Namibia", "Lesotho", "Eswatini"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly deaths")+theme_minimal()
```

#Western Africa 1

```{r}
weekly=coronavirus%>%filter(type=="death", country %in% c("Nigeria", "Ghana", "Côte d'Ivoire", "Niger", "Mali", "Burkina Faso", "Senegal", "Guinea"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly deaths")+theme_minimal()
```
#Western Africa 2

```{r}
weekly=coronavirus%>%filter(type=="death", country %in% c("Benin", "Togo", "Sierra Leone", "Liberia", "Mauritania", "Gambia", "Guinea-Bissau", "Cabo Verde"))%>%
  mutate(date=as.Date(date), weeks = floor_date(date, "weeks"))%>%group_by(country,weeks)%>%
  summarise(weekly_cases=sum(cases), avg_daily=round(sum(cases)/length(unique(date))))%>%rename(Country=country)

ggplot(data = weekly, aes(x=weeks, y=avg_daily, col=Country))+geom_line()+geom_point()+ylab("Weekly deaths")+theme_minimal()
```




```{r}
coronavirus %>%
filter(type == "death") %>%
group_by(country) %>%
summarise(total = sum(cases)) %>%
arrange(-total) %>%
print(n = 10)
```

#, "DR Congo", "Angola", "Cameroon", "Chad", "Congo", "Central African Republic", "Gabon", "Equatorial Guinea", "Sao Tome & Principe", "Egypt", "Sudan", "Algeria", "Morocco", "Tunisia", "Libya", "South Africa", "Botswana", "Namibia", "Lesotho", "Eswatini", "Nigeria", "Ghana", "Côte d'Ivoire", "Niger", "Mali", "Burkina Faso", "Senegal", "Guinea", "Benin", "Togo", "Sierra Leone", "Liberia", "Mauritania", "Gambia", "Guinea-Bissau", "Cabo Verde"

```{r}
Eastern_Africa1 = coronavirus %>% filter(country %in% c("Ethiopia", "Tanzania", "Kenya", "Uganda", "Mozambique", "Madagascar", "Malawi", "Zambia", "Somalia"), type=="death", date>as.Date("2020-02-1
5"))%>% group_by(type, country) %>% mutate(
cumDeaths = cumsum(cases))

ggplot(Eastern_Africa1, aes(x=date, y=cumDeaths)) +
geom_line(color="red") + facet_wrap(~country, ncol=3, scales="free_y") +
theme_classic() +
theme(axis.text.x = element_text(angle = 90)) +
labs(title="Accumulative Deaths by Covid-19 in Eastern Africa")
```



```{r}
Eastern_Africa2 = coronavirus %>% filter(country %in% c("Zimbabwe", "Rwanda", "Burundi", "South Sudan", "Eritrea", "Mauritius", "Djibouti", "Comoros", "Seychelles"), type=="death", date>as.Date("2020-02-1
5"))%>% group_by(type, country) %>% mutate(
cumDeaths = cumsum(cases))

ggplot(Eastern_Africa2, aes(x=date, y=cumDeaths)) +
geom_line(color="red") + facet_wrap(~country, ncol=3, scales="free_y") +
theme_classic() +
theme(axis.text.x = element_text(angle = 90)) + 
  labs(title="Accumulative Deaths by Covid-19 in Eastern Africa")
```









```{r}
Middle_Africa = coronavirus %>% filter(country %in% c("Democratic Republic of Congo", "Angola", "Cameroon", "Chad", "Congo", "Central African Republic", "Gabon", "Equatorial Guinea", "Sao Tome and Principe"), type=="death", date>as.Date("2020-02-1
5"))%>% group_by(type, country) %>% mutate(
cumDeaths = cumsum(cases))

ggplot(Middle_Africa, aes(x=date, y=cumDeaths)) +
geom_line(color="red") + facet_wrap(~country, ncol=3, scales="free_y") +
theme_classic() +
theme(axis.text.x = element_text(angle = 90)) + 
  labs(title="Accumulative Deaths by Covid-19 in Middle Africa")
```




```{r}
Northern_Africa = coronavirus %>% filter(country %in% c("Egypt", "Sudan", "Algeria", "Morocco", "Tunisia", "Libya"), type=="death", date>as.Date("2020-02-1
5"))%>% group_by(type, country) %>% mutate(
cumDeaths = cumsum(cases))

ggplot(Northern_Africa, aes(x=date, y=cumDeaths)) +
geom_line(color="red") + facet_wrap(~country, ncol=3, scales="free_y") +
theme_classic() +
theme(axis.text.x = element_text(angle = 90))+ 
  labs(title="Accumulative Deaths by Covid-19 in Northern Africa")
```



```{r}
Southern_Africa = coronavirus %>% filter(country %in% c("South Africa", "Botswana", "Namibia", "Lesotho", "Eswatini"), type=="death", date>as.Date("2020-02-1
5"))%>% group_by(type, country) %>% mutate(
cumDeaths = cumsum(cases))

ggplot(Southern_Africa, aes(x=date, y=cumDeaths)) +
geom_line(color="red") + facet_wrap(~country, ncol=3, scales="free_y") +
theme_classic() +
theme(axis.text.x = element_text(angle = 90))+ 
  labs(title="Accumulative Deaths by Covid-19 in Southern Africa")
```


```{r}
Western_Africa = coronavirus %>% filter(country %in% c("Nigeria", "Ghana", "Côte d'Ivoire", "Niger", "Mali", "Burkina Faso", "Senegal", "Guinea"), type=="death", date>as.Date("2020-02-1
5"))%>% group_by(type, country) %>% mutate(
cumDeaths = cumsum(cases))

ggplot(Western_Africa, aes(x=date, y=cumDeaths)) +
geom_line(color="red") + facet_wrap(~country, ncol=3, scales="free_y") +
theme_classic() +
theme(axis.text.x = element_text(angle = 90))+ 
  labs(title="Accumulative Deaths by Covid-19 in Western Africa")
```


```{r}
Western_Africa2 = coronavirus %>% filter(country %in% c("Benin", "Togo", "Sierra Leone", "Liberia", "Mauritania", "Gambia", "Guinea-Bissau", "Cabo Verde"), type=="death", date>as.Date("2020-02-1
5"))%>% group_by(type, country) %>% mutate(
cumDeaths = cumsum(cases))

ggplot(Western_Africa2, aes(x=date, y=cumDeaths)) +
geom_line(color="red") + facet_wrap(~country, ncol=3, scales="free_y") +
theme_classic() +
theme(axis.text.x = element_text(angle = 90)) + 
  labs(title="Accumulative Deaths by Covid-19 in Western Africa")
```




US	1123836			
Brazil	699276			
India	530779			
Russia	388478			
Mexico	333188			
United Kingdom	220721			
Peru	219539			
Italy	188322			
Germany

```{r}
Mostlethal = coronavirus %>% filter(country %in% c("US", "Brazil", "India", "Russia", "Mexico", "Peru"), type=="death", date>as.Date("2020-02-1
5"))%>% group_by(type, country) %>% mutate(
cumDeaths = cumsum(cases))

ggplot(Mostlethal, aes(x=date, y=cumDeaths)) +
geom_line(color="red") + facet_wrap(~country, ncol=3, scales="free_y") +
theme_classic() +
theme(axis.text.x = element_text(angle = 90)) + 
  labs(title="Accumulative Deaths by Covid-19 in Most Lethal countries")
```